Design and Implementation of Probabilistic Programming Language Anglican

### 2016-08-15 paper/paper.tex
rewriting in progress

% Authors: use changebars to mark changed regions
% \cbstart
% ...
% \cbend

@@

An implementation of Anglican must therefore address three issues:
\begin{itemize}
    \item the Clojure syntax to introduce probabilistic Anglican
        code inside Clojure modules;
    \item source-to-source transformation of Anglican programs
        into Clojure, so that probabilistic execution becomes
        possible;
    \item algorithms which run Clojure code, obtained by
        transforming Anglican programs, according to the
        probabilistic operational semantics.
\end{itemize}

@@

\hy{Great introduction. It very nicely explains why Anglican
is born and how it is implemented. If I act as a really picky
reviewer or a reader, I would also expect the instroduction
to help me grasp in a high level what new insights or scientific problems 
this development of Anglican have resulted in or solved so far.}

@@

Anglican can be mostly implemented as a regular programming
language, except for the handling of these checkpoints.
Depending on the \textit{inference algorithm}, \texttt{sample}
and \texttt{observe} may result in implicit input/output
operations and control changes. For example, \texttt{observe} in
particle filtering inference algorithms \cite{WVM14} is a non-deterministic
control statement at which a particle can be either replicated
or terminated. Similarly, in Metropolis-Hastings \cite{WSG11},
\texttt{sample} is both an input and a non-deterministic control
statement (with delayed effect), eventually affecting acceptance
or rejection of a sample.

\hy{What do you mean by input in ``both an input..''?}

>>>>

\cbcolor{red}
% Authors: use changebars to mark changed regions
% \cbstart
% ...
% \cbend

@@

An implementation of Anglican must therefore address three issues:
\begin{itemize}
    \item the Clojure syntax to introduce probabilistic Anglican
        code inside Clojure modules;
    \item source-to-source transformation of Anglican programs
        into Clojure, so that probabilistic execution becomes
        possible;
    \item algorithms which run Clojure code, obtained by
        transforming Anglican programs, according to the
        probabilistic operational semantics.
\end{itemize}
\cbstart
Execution of probabilistic programs by inference algorithms is
different from execution of deterministic programs.  A
probabilistic program is executed multiple times, often hundreds
of thousands or even millions of times for a single inference
task. Random choices may affect which parts of the program code
are executed and how often. Many inference algorithms require
re-running the program multiple times partially, from a certain
point on.  Different executions may employ different random
choices, however for efficient inference a correspondence
between random choices in different executions should be
maintained. These are just some of the challenges which were
faced and solved during development of Anglican.
\cbend

@@

% \hy{Great introduction. It very nicely explains why Anglican
% is born and how it is implemented. If I act as a really picky
% reviewer or a reader, I would also expect the introduction
% to help me grasp in a high level what new insights or scientific problems 
% this development of Anglican have resulted in or solved so far.}

@@

Anglican can be mostly implemented as a regular programming
language, except for the handling of these checkpoints.
Depending on the \textit{inference algorithm}, \texttt{sample}
and \texttt{observe} may result in implicit input/output
operations and control changes. For example, \texttt{observe} in
particle filtering inference algorithms \cite{WVM14} is a non-deterministic
control statement at which a particle can be either replicated
or terminated. Similarly, in Metropolis-Hastings \cite{WSG11},
\texttt{sample} is both an input
\cbstart
statement which `reads' values from a random source,
\cbend
and a non-deterministic control statement
(with delayed effect), eventually affecting acceptance or
rejection of a sample.

%\hy{What do you mean by input in ``both an input..''?}
